\bf Estimating the number of clusters using cross-validation

### 2016-06-20 cvclust.tex
fix a few typos

\if1\blind
{
  \title{\bf Estimating the number of clusters using cross-validation}
  \author{Wei Fu and Patrick O. Perry \\
  Stern School of Business, New York University}
  \maketitle
} \fi

@@

The transformation used in step~\ref{step:transform} uses a random orthogonal
matrix~$Q$, which gets applied to the rows of $\dataX$ after 
multiplying by the estimate of $\Sigma^{-1/2}$. We use this random orthogonal
matrix for two reasons. First, it ensures that in expectation, each transformed cluster mean
$Q \Lambda^{-1/2} \Gamma^\T \mu(g)$ for $g = 1, \dotsc, k$ is uniformly
spread across all $P$ features. This ensures that the self-consistency
conditions on the cluster centers enumerated in Proposition~\ref{prop:self-consistency}
are likely to hold. The second reason for multiplying by $Q$ is to spread any
remaining correlation in the noise evenly (in expectation) across all
dimensions. The latter effect follows since if $Z$ is a random vector with
covariance matrix $\Theta$, then $Q Z$ has covariance matrix $Q \Theta Q^\T$,
which has expectation $E(Q \Theta Q^\T) = \tr(\Theta) I$.

@@

\begin{table}
\centering
\captionsetup{justification=centering}
\caption{\label{tab:benchmark} Number of clusters selected on benchmark datasets}
\begin{tabular}{lccc}
\toprule                  
         & \multicolumn{3}{c}{Dataset} \\
\cmidrule(l){2-4}
Method   & Congress Voting & Breast Cancer & Brain Tumours \\
\midrule                    
Gabriel  & $2$             & $3$           & $5$  \\    
Gabriel (corr.~correct) & $2$ & $2$           & $5$  \\  
Wold     & $2$             & $3$           & $4$  \\
Gap      & $8$             & $10$          & $10$ \\   
BIC      & $2$         & $5$           & $2$  \\   
CH       & $2$             & $2$           & $2$  \\
Hartigan & $3$             & $3$           & $4$  \\
Jump     & $10$            & $9$           & $1$  \\   
PS       & $2$  & $2$           & $1$  \\
Stab.    & $2$  & $2$           & $7$  \\ 
\midrule 
Ground Truth & $2$         & $2$ or $3$    & $5$  \\
\bottomrule
\end{tabular}
\end{table}

@@

From Table \ref{table:enrich}, we can see that Cluster~1 is enriched with
genes that somatize cell stress, such as oxidative heat-induce proteins.
Cluster~2 contains genes that govern mitochondrial translation and
mitochondrion organization. Cluster~3, the first period cluster, contains cell
cycle genes related to budding and cell polarity, along with genes that govern
RNA processing and transcription. Cluster~4 contains genes related to
cytoplasmic translation and genes encoding ribosomes. Cluster~5, the second
periodic cluster, contains genes that participate cell-cycle processes, along
with DNA replication and DNA repair.

>>>>

\if1\blind
{
  \title{\bf Estimating the number of clusters using cross-validation}
  \author{Wei Fu and Patrick O.\ Perry \\
  Stern School of Business, New York University}
  \maketitle
} \fi

@@

The transformation used in step~\ref{step:transform} uses a random orthogonal
matrix~$Q$, which gets applied to the rows of $\dataX$ after 
multiplying by the estimate of $\Sigma^{-1/2}$. We use this random orthogonal
matrix for two reasons. First, it ensures that in expectation, each transformed cluster mean
$Q \Lambda^{-1/2} \Gamma^\T \mu(g)$ for $g = 1, \dotsc, k$ is uniformly
spread across all $P$ features. This ensures that the self-consistency
conditions on the cluster centers enumerated in Proposition~\ref{prop:self-consistency}
are likely to hold. The second reason for multiplying by $Q$ is to spread any
remaining correlation in the noise evenly (in expectation) across all
dimensions. The latter effect follows since if $Z$ is a random vector with
covariance matrix $\Theta$, then $Q Z$ has covariance matrix $Q \Theta Q^\T$,
which has expectation $\E(Q \Theta Q^\T) = \tr(\Theta) I$.

@@

\begin{table}
\centering
\captionsetup{justification=centering}
\caption{\label{tab:benchmark} Number of clusters selected on benchmark datasets}
\begin{tabular}{lccc}
\toprule                  
         & \multicolumn{3}{c}{Dataset} \\
\cmidrule(l){2-4}
Method   & Congress Voting & Breast Cancer & Brain Tumors \\
\midrule                    
Gabriel  & $2$             & $3$           & $5$  \\    
Gabriel (corr.~correct) & $2$ & $2$           & $5$  \\  
Wold     & $2$             & $3$           & $4$  \\
Gap      & $8$             & $10$          & $10$ \\   
BIC      & $2$         & $5$           & $2$  \\   
CH       & $2$             & $2$           & $2$  \\
Hartigan & $3$             & $3$           & $4$  \\
Jump     & $10$            & $9$           & $1$  \\   
PS       & $2$  & $2$           & $1$  \\
Stab.    & $2$  & $2$           & $7$  \\ 
\midrule 
Ground Truth & $2$         & $2$ or $3$    & $5$  \\
\bottomrule
\end{tabular}
\end{table}

@@

From Table \ref{table:enrich}, we can see that Cluster~1 is enriched with
genes that somatize cell stress, such as oxidative heat-induce proteins.
Cluster~2 contains genes that govern mitochondrial translation and
mitochondrion organization. Cluster~3, the first period cluster, contains cell
cycle genes related to budding and cell polarity, along with genes that govern
RNA processing and transcription. Cluster~4 contains genes related to
cytoplasmic translation and genes encoding ribosomes. Cluster~5, the second
periodic cluster, contains genes that participate in cell-cycle processes, along
with DNA replication and DNA repair.
